---
title: "Recode and Map Geo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recode and Map Geo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide explains how to create geo table to recode and map geo codes. This
file will be saved in database file specified in **orgdata.geo** option. See
`getOptions("orgdata.geo")`

# Recode Geo Codes

The function to create table for recoding geo codes is `geo_recode()`. The table
will consist of previous codes and their new codes whenever available. For
instance, to get all codes for *kommuner* from 2017 to 2022. The table will list
all the codes for *kommuner* since 2017 in columname `oldCode` while their current
codes ie. 2022, is in colum `currentCode`. Any geo codes that have changed since
2017 will have their new codes in `currentCode` column. For geo codes that have
not changed since 2017 will have the value as `NA` in the `oldCode` column.

Use argument `write = TRUE` to save it in the database. If the table already
exists in the database and you just want to add a new dataset to the table, use
argument `append = TRUE`. The table in the database will be called *kommune2022*
if the data consist of geo codes up to 2022, ie. argument `to = 2022` and `typ =
kommune`.

However in most cases you will only need to use argument `write = TRUE`. Check
the documentation for possible arguments with `help("geo_recode")`.

```{r geo-recode, eval = FALSE}
library(orgdata)
dt <- geo_recode(type = "kommune", from = 2018, to = 2022)
dt[1:5]
```

```r 
oldCode              oldName currentCode           newName changeOccurred      batch
1:    <NA>              <NA>        0301              Oslo           2022 2022-03-15
2:    <NA>              <NA>        1101         Eigersund           2022 2022-03-15
3:    1141            Finnøy        1103         Stavanger           2020 2022-03-15
4:    1142          Rennesøy        1103         Stavanger           2020 2022-03-15
5:    <NA>              <NA>        1106         Haugesund           2022 2022-03-15
```

# Mapping Geo Codes

The function to create geo mapping is `geo_map()`. The default table name to
be created in the database is *tblGeo*. It consists of all the geo granularities
(`levels`) of the selected year. The lowest level is enumeration areas codes or
*grunnkrets*.

Arguments `write` and `append` can be used when calling `geo_map()` function.
However in most cases, you will only need to use `append = TRUE`, ie. to append
the data to the existing *tblGeo* in the database.

```{r geo-map, eval = FALSE}
DT <- geo_map(year = 2022)
DT[sample(1:.N, 5)]
```

```r
code                      name validTo      level grunnkrets kommune fylke bydel      batch
1: 30470104 Søndre Simostranda    2022 grunnkrets   30470104    3047    30  <NA> 2022-03-15
2: 11140110           Vikeså 2    2022 grunnkrets   11140110    1114    11  <NA> 2022-03-15
3: 30020408           Ørehavna    2022 grunnkrets   30020408    3002    30  <NA> 2022-03-15
4: 46350304          Hjartholm    2022 grunnkrets   46350304    4635    46  <NA> 2022-03-15
5: 42160109     Birkeland Nord    2022 grunnkrets   42160109    4216    42  <NA> 2022-03-15
```


# Merge Geo Codes

Both `geo_recode()` and `geo_map()` fetch the data via API available from
SSB. To merge geo codes that aren't available via API can be done with the
function `geo_merge()`.

The data to merge can be in any format that is accepted by `read_file()`
function. The new data must have column id to merge into the database *tblGeo*
and column with the new geo codes to be added to. For instance if a new dataset
from 2022 has `levekaar` codes in columname `code_levekaar` and these `levekaar`
codes are derived from `bydel` codes, then the specification for the arguments
will be as the example below.

Please read the documentation for details on arguments that can be used with
`help("geo_merge")` or `?geo_merge`.

```{r geo-merge, eval = FALSE}
dd <- geo_merge(id.table = "bydel",
                id.file = "var01",
                geo.col = "code_levekaar",
                geo.level = "levekaar",
                file = "C:/Users/ybka/data/levekaar.csv",
                year = 2022)

```

