---
title: "Recode and Map Geo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recode and Map Geo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide explains how to create geo table to recode and map geo codes. This
file will be saved in **orgdata.geo** database. See `getOptions("orgdata.geo")`

# Recode Geo Codes

The function to create table for recoding geo is `geo_recode()`. The table
will consist of previous codes and their new codes whenever available. For
instance, to get all codes for *kommuner* from 2017 to 2022. The table will list
all the codes for *kommuner* since 2017 in columname `oldCode` while their current
codes ie. 2022, is in colum `currentCode`. Any geo codes that have changed since
2017 will have their new codes in `currentCode` column. For geo codes that have
not changed since 2017 will have the value as `NA` in the `oldCode` column.

Use argument `write = TRUE` to save it in the database. If the table already
exists in the database and you just want to add a new dataset to the table, use
argument `append = TRUE`. The table in the database will be called *kommune2022*
if the data consist of geo codes up to 2022, ie. argument `to = 2022` and `typ =
kommune`.

However in most cases you will only need to use argument `write = TRUE`. Check
the documentation with `help("geo_recode")`.

```{r geo-recode, include = TRUE }
library(orgdata)
dt <- geo_recode(type = "kommune", from = 2018, to = 2022)
dt[1:5]
```

# Mapping Geo Codes

The function to create geo mapping is `geo_levels()`. The default table name to
be created in the database is *tblGeo*. It consists of all the geo granularities
(`levels`) of the selected year. The lowest level is enumeration areas codes or
*grunnkrets*.

Arguments `write` and `append` can be use when calling `geo_levels()` function.
However in most cases, you will only need to use `append = TRUE`, ie. to append
the data to the existing *tblGeo* in the database.

```{r geo-map, include = TRUE }
DT <- geo_levels(year = 2022)
DT[sample(1:.N, 5)]
```

# Merge Geo Codes

Both `geo_recode()` and `geo_levels()` fetch the data via API available from
SSB. To merge geo codes that aren't available via API can be done with the
function `geo_merge()`.

The data to merge can be in any format that is accepted by `read_file()`
function. The data must consist of column id to merge to the database *tblGeo*
and column with the new geo codes to be added to.

Below is an example on how the function can be specified. Please read the
documentation for type of arguments that can be accepted.

```{r geo-merge, eval = FALSE}
dd <- geo_merge(id.table = "kommune",
                id.file = "var01",
                geo.col = "code_levekaar",
                geo.level = "levekaar",
                file = "C:/Users/ybka/data/levekaar.csv",
                year = 2022)

```

